A free/open-source Kazakh-Tatar machine translation system

### 2013-04-06 paper/paper.tex
beginnings of comparative grammar

Both Tatar and Kazakh belong to the Kypchak (or Northwestern) group of Turkic
languages. Kazakh is primarily spoken in Kazakhstan, where it is the national
language. Large groups of native speakers also exist in China, neigbouring
Central-Eurasian republics, and Mongolia. The total number of speakers is at least
10 million people. Tatar is a Turkic language spoken in and around Tatarstan by
approximately 6 million people. It is co-official with Russian in Tatarstan --
a republic within Russia.  A majority of native speakers of both languages are bilingual in Russian.

%FIXME: Contrastive grammar

>>>>

Both Tatar and Kazakh belong to the Kypchak (or Northwestern) group of Turkic
languages.  They share some level of mutual intelligibility to naïve native speakers, though this is somewhat limited, especially regarding the written languages.

Kazakh is primarily spoken in Kazakhstan, where it is the national
language, sharing official status with Russian as the FIXME language.
Large groups of native speakers also exist in China, neigbouring
Central-Eurasian republics, and Mongolia. The total number of speakers is at least 10 million people.

Tatar is spoken in and around Tatarstan by
approximately 6 million people. It is co-official with Russian in Tatarstan --
a republic within Russia.  A majority of native speakers of both languages are bilingual in Russian.

%FIXME: Contrastive grammar
\subsection{Phonological differences}
%a generalised summary, 2 or 3 small specific examples
As closely related languages, Kazakh and Tatar share many phonological processes, including front-back vowel harmony systems, consonant voicing assimilation, and even a typologically rare consonantal nasal harmony system.  However, the differing details of these processes and the existence of processes unique to each language render Kazakh and Tatar fairly different.  For example, Kazakh has a ubiquitous system of desonorisation of the initial sonorants found in many common morphemes.  Furthermore, Tatar has nasal assimilation of the initial /l/ of the plural-suffix.

\subsection{Orthographic differences}
a generalised summary, 1 or 2 small specific examples
\subsection{Morphotactic differences}
2 or 3 specific examples
\subsection{Lexical and morphological differences}
2 or 3 specific examples
\subsection{Syntactic differences}
2 or 3 specific examples
> ұна- vs.\ ярат- (does this count as syntactic?)
